What words ought to exist? \\ {\normalsize Coining with coinduction}

### 2011-03-27 wishlist/paper.tex
notes about backformation, which is not actually implemented yet
more progres on paper. compute bigram frequencies for italian

My website ``{\sf snoot.org}'' has a number of games on it, including
a Scrabble clone called Scribble\footnote{{\tt
    http://snoot.org/toys/scribble/}} and Boggle clone called
Muddle.\!\footnote{{\tt http://snoot.org/toys/muddle/}} This website
has been running for almost ten years, comprising over 150,000
Scribble games totaling 3.8 million words placed and 628,000 Muddle
games with over 10 million words found. During each game, players
repeatedly attempt to play words that aren't real. The computer
rebukes them, but hope really springs eternal with these people. It's
like they truly deeply wish to break out of the shackles of the
Official Scrabble Players Dictionary.\!\footnote{For the analyses in
  this paper that depend on a list of legal words, I actually use a
  modified version of SOWPODS, which is the tournament list used in
  Australia and the UK, and significantly more permissive than the US
  Tournament Word List. Though the modified version is non-canonical,
  I stuck with it because it's what's been in use on the site for
  ten years.} So the first approach to determining what
words ought to exist is to analyze the words that people tried to
play, in order to try to extract the essence of word-yearning.

@@

\begin{figure}
\includegraphics[width=\linewidth]{wishlist-cdf}
\caption{Distribution of word frequency. Approximately 25,000
  different words (y axis) were issued 55 times or fewer (x axis).}
\label{fig:gamedistribution}
\end{figure}

@@

\section{Statistical Models}

@@

% XXX improve layout
\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
0.049901545  & s     &   0.001768358  & y   \\
0.017526947  & d     &   0.001761301  & p   \\
0.009521421  & g     &   0.001653041  & a   \\
0.005506083  & c     &   0.001652864  & n   \\
0.004376836  & r     &   0.001563810  & ps  \\
0.004218223  & t     &   0.001383503  & ms  \\
0.004088377  & e     &   0.001371545  & ts  \\
0.003560976  & m     &   0.001327023  & ds  \\
0.003248559  & ss    &   0.001188715  & hy  \\
0.002001529  & rs    &   0.001166457  & k   \\
0.001958674  & h     &   0.001160619  & ng  \\
0.001873761  & l     &   0.001142785  & ly  \\
\end{tabular}
\end{center}

@@

\begin{verbatim}
Markov chain with n=2.
100 top paths:
0.007092576 ing
0.002483309 ses
0.001694275 des
0.001548279 nes
0.001403047 sts
0.001314279 se
0.001287353 ings
0.001262904 ded
0.001170356 cal
0.001101124 le
0.001078157 der
0.001070769 ove
0.001016514 gly
0.000882527 hy
0.000853499 ung
0.000838535 cy
0.000816757 pres
0.000802384 pers
0.000787731 ps
0.000771953 ders
0.000746855 ly
0.000732157 ines
0.000709460 ve
0.000650273 sm
0.000650062 ty
0.000642246 pred
0.000627432 ble
0.000622587 hed
0.000603012 ter
0.000587263 gy
0.000564629 sms
0.000549088 py
0.000545497 ans
0.000541832 ber
0.000531745 ce
0.000489139 ang
0.000487316 unt
0.000482908 snes
\end{verbatim}

\begin{verbatim}
Markov chain with n=3.
100 top paths:
0.001093377 des
0.000789743 pers
0.000767821 cal
0.000622381 pres
0.000452415 nons
0.000449855 ress
0.000426892 ing
0.000404454 pred
0.000387570 ent
0.000366753 dist
0.000357115 ble
0.000353576 ches
0.000348642 gly
0.000345564 inted
0.000340691 dists
0.000339750 lity
0.000336468 und
0.000331914 dises
0.000331389 int
0.000320104 ints
0.000291846 ching
0.000284350 tely
0.000284228 dise
0.000279669 ster
0.000279431 ves
0.000276676 stic
0.000272655 tric
0.000263922 cally
0.000261415 ter
0.000260898 ents
0.000258270 rous
0.000253990 dising
0.000245834 phy
0.000244272 dised
0.000240505 hous
0.000239726 resses
0.000237100 sters
0.000236753 ove
0.000230976 dism
0.000227701 ched
0.000225104 fing
0.000222993 ence
0.000221624 ters
0.000219511 disms
0.000219246 unders
\end{verbatim}

\begin{verbatim}
Markov chain with n=4.
100 top paths:
0.000454575 unders
0.000341807 dising
0.000293774 pers
0.000285248 cally
0.000232842 inted
0.000204856 heter
0.000198642 tric
0.000188363 ster
0.000185238 hier
0.000183216 unded
0.000174876 heters
0.000169220 sters
0.000153800 stic
0.000142753 pering
0.000132595 dises
0.000131040 ching
0.000129387 shing
0.000125442 dest
0.000115038 teless
0.000113312 resis
0.000112909 dist
0.000112793 dists
0.000112397 mising
0.000109589 sation
0.000109555 fing
0.000108931 dised
0.000107632 ches
0.000102999 cated
0.000101327 dise
0.000097730 sations
0.000096676 inding
0.000095270 nons
0.000093852 lary
0.000093357 ency
0.000091199 ther
0.000090608 ories
0.000090405 alled
0.000088604 apped
0.000088235 hous
0.000087780 trous
0.000082438 undered
0.000081713 noning
0.000079613 dism
0.000079071 thers
0.000078380 ented
0.000076765 irred
0.000076363 disms
0.000074970 geness
0.000074910 veness
0.000071079 vities
0.000070308 mised
0.000068585 ques
0.000068096 unding
0.000067618 sher
0.000067429 peries
\end{verbatim}

\begin{verbatim}
n=5
GetTempFileName failed with error 5
\end{verbatim}

@@

% laplace smoothing?
% TODO: the empty string?

TODO: most likely words in italian, too.

% word list from Italian Scrabble, which I bet is called
% {Scrabbizzimo!}

@@

@@

Rob: etsy, nuuog
Chris: tapping on the bisphenol-A, nurm
David: wafflucinations
Lea: hnfff sound
Reed: pansepticon
Jessica: gruntle (like not disgruntled)

In order to not reprint everyone's bullshit---but not introduce bias by selectively removing data---I discarded random subsets of the data until it did not contain bullshit any more.

@@

\section{Recommendations}

@@

>>>>

My website ``{\sf snoot.org}'' has a number of games on it, including
a Scrabble clone called Scribble\footnote{{\tt
    http://snoot.org/toys/scribble/}} and Boggle clone called
Muddle.\!\footnote{{\tt http://snoot.org/toys/muddle/}} This website
has been running for almost ten years, comprising over 150,000
Scribble games totaling 3.8 million words placed and 628,000 Muddle
games with over 10 million words found. During each game, players
repeatedly attempt to play words that aren't real. The computer
rebukes them, but hope really springs eternal with these people. It's
like they truly deeply wish to break out of the shackles of the
Official Scrabble Players Dictionary.\!\footnote{For the analyses in
  this section that depend on a list of legal words, I actually use a
  modified version of SOWPODS, which is the tournament list used in
  Australia and the UK, and significantly more permissive than the US
  Tournament Word List. Though the modified version is non-canonical,
  I stuck with it because it's what's been in use on the site for
  ten years.} So the first approach to determining what
words ought to exist is to analyze the words that people tried to
play, in order to try to extract the essence of word-yearning.

@@

\begin{figure}
\includegraphics[width=\linewidth]{wishlist-cdf}
\vspace{-0.1in}\caption{Cumulative distribution of word frequency.
  Approximately 25,000 different words (y axis) were issued 55 times
  or fewer (x axis). The ``total'' area does not appear much larger
  than its components because this is a log-log plot.}
\label{fig:gamedistribution}
\end{figure}

@@

\section{Statistical models}

@@

% XXX improve layout
\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
4.99\%  & s     &   0.17\%  & y   \\
1.75\%  & d     &   0.17\%  & p   \\
0.95\%  & g     &   0.16\%  & a   \\
0.55\%  & c     &   0.16\%  & n   \\
0.43\%  & r     &   0.15\%  & ps  \\
0.42\%  & t     &   0.13\%  & ms  \\
0.40\%  & e     &   0.13\%  & ts  \\
0.35\%  & m     &   0.13\%  & ds  \\
0.32\%  & ss    &   0.11\%  & hy  \\
0.20\%  & rs    &   0.11\%  & k   \\
0.19\%  & h     &   0.11\%  & ng  \\
0.18\%  & l     &   0.11\%  & ly  \\
\end{tabular}
\end{center}

@@

\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
\multicolumn{4}{c}{Markov chain with $n=2$.} \\
\hline
.709\% & ing     &   .110\% & le    \\
.248\% & ses     &   .107\% & der   \\
.169\% & des     &   .107\% & ove   \\
.154\% & nes     &   .101\% & gly   \\
.140\% & sts     &   .088\% & hy    \\
.131\% & se      &   .085\% & ung   \\
.128\% & ings    &   .083\% & cy    \\
.126\% & ded     &   .081\% & pres  \\
.117\% & cal     &   .080\% & pers  \\
\end{tabular}
\end{center}

These are even, like, pronounceable. The best news is that they keep
getting better the more history we keep:

\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
\multicolumn{4}{c}{Markov chain with $n=3$.} \\
\hline
.109\% & des     &   .038\%  & ent     \\
.078\% & pers    &   .036\%  & dist    \\
.076\% & cal     &   .035\%  & ble     \\
.062\% & pres    &   .035\%  & ches    \\
.045\% & nons    &   .034\%  & gly     \\
.044\% & ress    &   .034\%  & inted   \\
.042\% & ing     &   .034\%  & dists   \\
.040\% & pred    &   .033\%  & lity    \\
\end{tabular}
\end{center}

\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
\multicolumn{4}{c}{Markov chain with $n=4$.} \\
\hline
.045\% & unders  &   .017\% & heters  \\
.034\% & dising  &   .016\% & sters   \\
.029\% & pers    &   .015\% & stic    \\
.028\% & cally   &   .014\% & pering  \\
.023\% & inted   &   .013\% & dises   \\
.020\% & heter   &   .013\% & ching   \\
.019\% & tric    &   .012\% & shing   \\
.018\% & ster    &   .012\% & dest    \\
.018\% & hier    &   .011\% & teless  \\
.018\% & unded   &   .011\% & resis   \\
\end{tabular}
\end{center}

\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
\multicolumn{4}{c}{Markov chain with $n=5$.} \\
\hline
\multicolumn{4}{l}{{\tt GetTempFileName failed with error 5}} \\
\end{tabular}
\end{center}

With four letters of history, the words produced are quite good! (The
results at $n=5$ are somewhat disappointing since the program crashes
from running out of memory. The table at $n=5$ would have over 481
million entries.) Many of these seem like real words. Some even
suggest meaning because they contain common morphemes. To make the
case that these are not just real-looking words but characteristic of
the English language, compare the results of the same algorithm on the
dictionary from the Italian language edition of Scrabble, which is
probably called {\it Scrabblizzimo!} (Figure~\ref{fig:italian}).
Italian is lexicographically a more compact language than English
(Figure~\ref{fig:italianbigrams}); there are only 21 letters (outside
of occasional interlopers in loan words like {\it jeans} and {\it
  taxi}). Moreover, even though the dictionary contains 585,000 words
(twice as many as English), the probabilities of observing these
non-words are much higher than the most likely English ones.

\begin{figure}
\includegraphics[width=\linewidth]{italiancoin1}
\caption{Markov chain for the Italian language. Again darker cells
  indicate higher probability. Italian has more lexicographic
  structure recognizable from bigraphs than English does: Note that
  the extremely rare letters ``j'', ``k'', ``q'', ``w'', ``x'', and
  ``y'' have almost empty rows. ``z'' very frequently follows ``z'',
  as in {\it pizza}. Words almost always end in a vowel.}
\label{fig:italianbigrams}
\end{figure}

\begin{figure}
\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
.137\%  &  ammo       &   .026\%  &  rino      \\
.071\%  &  rice       &   .025\%  &  diste     \\
.061\%  &  rico       &   .024\%  &  risti     \\
.055\%  &  este       &   .023\%  &  disci     \\
.053\%  &  scono      &   .022\%  &  riasse    \\
.049\%  &  immo       &   .022\%  &  riassi    \\
.047\%  &  assero     &   .021\%  &  cate      \\
.047\%  &  scano      &   .019\%  &  rite      \\
.038\%  &  rammo      &   .019\%  &  cando     \\
.034\%  &  cata       &   .018\%  &  riassero  \\
.034\%  &  assimo     &   .018\%  &  riassimo  \\
.032\%  &  riate      &   .018\%  &  dete      \\
.032\%  &  disce      &   .018\%  &  disca     \\
.030\%  &  esti       &   .017\%  &  risca     \\
.029\%  &  rica       &   .017\%  &  cente     \\
.028\%  &  endo       &   .016\%  &  acci      \\
.027\%  &  dissimo    &   .015\%  &  centi     \\
.026\%  &  rici       &   .015\%  &  girono    \\
\end{tabular}
\end{center}
\caption{Most probable words induced by the Markov Markov chain for
  the Italian language ($n=4$).}
\label{fig:italian}
\end{figure}

\subsection{Usage-weighted methods}

One criticism of this approach is that it considers every word in the
word list to be equally important.\!\footnote{In fact, the common part
  of words with many different conjugations is in essence counted many
  times. This means {\it ornithology} in its six different forms
  contributes six times as much to our model as the word {\it the}!} I
object on the philosophical grounds that some words that already exist
{\em ought to exist more} than other words that already exist. For
example, {\it congenital} is a much nicer word than the plain ugly
{\it congenial}, and is reflected by the fact that {\it congenial} is
used five times more frequently than {\it
  congenial}.\!\footnote{14,200,000 times to 2,820,000, on the
  Internet, according to Google.} In this section, we produce Markov
models of words weighted by the frequency with which people tend to
use them. This is just a simple matter of training the model on some
natural language corpus (with many occurrences of each word, or no
occurrences of unpopular words) rather than a flat list of all alleged
words.

Since the best research is intensely navel-gazing, I started by 
analyzing a corpus of my own writing, specifically my Facebook status
updates since March~2006. There were 1,386 status updates containing
such jems as ``Tom Murphy VII thinks mathfrak is straight ballin''
and ``Tom Murphy VII global L\_50 reused for unused\_36166!!''. The
most likely words with $n=4$:

\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
.252\% &  pittsburgh   & .097\% &  can't    \\   
.209\% &  steelers     & .083\% &  i'm      \\   
.209\% &  sfo          & .083\% &  icfp     \\   
.195\% &  it's         & .083\% &  app      \\   
.125\% &  bdl          & .069\% &  x        \\   
.111\% &  sigbovik     & .069\% &  drunj    \\   
.109\% &  facebook     & .069\% &  g        \\   
.097\% &  mic          & .061\% &  ther     \\   
.097\% &  s            & .055\% &  doesn't  \\
\end{tabular}
\end{center}

This is the worst. Not only does it contain loads of one-letter words that
we have already determined are verboten\footnote{Note that since $n=4$ these
words have to actually appear in status updates to have nonzero probability for
this list. ``g'' is explained by frequent occurrences of ``e.g.'', for example.},
but the rest are just non-words that I tend to use like the names of cities,
prestigious conferences, or IATA airport codes.

@@

% laplace smoothing?

@@

\section{Special cases}

The empty string??
% TODO: the empty string?

\section{Backformation} \label{sec:backformation}

\begin{verbatim}
Best fixes...
1.000000000 -zzyingly    1/1
1.000000000 -zzying      1/1
1.000000000 -zzuolanas   1/1
1.000000000 -zzuolana    1/1
1.000000000 -zzotints    1/1
1.000000000 -zzotintos   1/1
1.000000000 -zzotinto    1/1
1.000000000 -zzotinting  1/1
1.000000000 -zzotinters  1/1
1.000000000 -zzotinter   1/1
1.000000000 -zzotinted   1/1
1.000000000 -zzotint     1/1
1.000000000 -zzonites    1/1
1.000000000 -zzonite     1/1
1.000000000 -zzoners     1/1
1.000000000 -zzoner      1/1
1.000000000 -zzolans     1/1
1.000000000 -zzolanic    1/1
1.000000000 -zzolan      1/1
1.000000000 -zzmen       1/1
1.000000000 -zzman       1/1
1.000000000 -zzlike      1/1
1.000000000 -zzini       1/1
1.000000000 -zzicatos    1/1
1.000000000 -zzicato     1/1
1.000000000 -zzicati     1/1
1.000000000 -zzian       1/1
\end{verbatim}
1/1 is not really much support.

Ones that are only seen n/1 time never apply (because ...). Filtered:

(actually we should probably be using binomial likelihood)

\begin{verbatim}
Best fixes...
1.000000000 -zzazzes     3/3
1.000000000 -zzazz       3/3
1.000000000 -zzans       3/3
1.000000000 -zzanim      2/2
1.000000000 -zzan        3/3
1.000000000 -zygotic     3/3
1.000000000 -zygotes     2/2
1.000000000 -zygote      2/2
1.000000000 -zygosity    2/2
1.000000000 -zygosis     2/2
1.000000000 -zpas        2/2
1.000000000 -zpahs       2/2
1.000000000 -zpah        2/2
1.000000000 -zpa         2/2
1.000000000 -zooks       2/2
1.000000000 -zonites     3/3
1.000000000 -zonite      3/3
1.000000000 -zonians     2/2
1.000000000 -zonian      2/2
1.000000000 -zists       2/2
1.000000000 -zist        2/2
\end{verbatim}

Not much better!

Only ones with denominator 10+ times.
\begin{verbatim}
Best fixes...
1.000000000 -wrought    10/10
1.000000000 -writings   12/12
1.000000000 -wraps      10/10
1.000000000 -wrap       11/11
1.000000000 -worms      69/69
1.000000000 -worm       69/69
1.000000000 -working    21/21
1.000000000 -worked     14/14
1.000000000 -winds      17/17
1.000000000 -wind       18/18
1.000000000 -wide       12/12
1.000000000 -weights    19/19
1.000000000 -weight     22/22
1.000000000 -waxes      12/12
1.000000000 -wax        12/12
1.000000000 -waters     37/37
1.000000000 -water      41/41
1.000000000 -walking    11/11
1.000000000 -walkers    17/17
1.000000000 -walker     17/17
1.000000000 -valuing    10/10
1.000000000 -valued     11/11
1.000000000 -towns      11/11
1.000000000 -tight      15/15
1.000000000 -thought    12/12
1.000000000 -thinks     11/11
\end{verbatim}

Much better! These are still sort of degenerate and won't help
us find new words because they ALWAYS apply!

Instead anything with p strictly less than 1.
\begin{verbatim}
Best fixes...
0.985714286 -makers     69/70
0.985714286 -maker      69/70
0.983606557 -wood       120/122
0.983471074 -woods      119/121
0.982758621 -down       57/58
0.982658960 -works      170/173
0.981818182 -houses     108/110
0.981818182 -house      108/110
0.981132075 kilo-       52/53
0.980752406 -less       1121/1143
0.980743395 over-       2190/2233
0.980000000 -books      49/50
0.980000000 -book       49/50
0.979591837 -proof      48/49
0.979310345 -lessnesses 142/145
0.979069767 -ships      421/430
0.978723404 -lessness   184/188
0.978723404 -board      138/141
0.978494624 -woman      91/93
0.978021978 -women      89/91
0.977528090 -ship       435/445
0.977272727 -manship    43/44
0.976744186 -weeds      84/86
0.976470588 after-      83/85
0.976190476 -manships   41/42
0.976190476 -making     41/42
0.976190476 -craft      41/42
0.976190476 -boats      41/42
0.976190476 -boat       41/42
\end{verbatim}

Wow, now we're talking!

@@

I asked a few people who happened to be around, ``What word ought to
exist?'' Most people did not know what to make of this question, and
also, because people seem to revel in the opportunity to get (well
deserved) revenge on me by being disruptive trolls, many of the
answers were designed to be unusable. In order to not reprint
everyone's bullshit---but not introduce bias by selectively removing
data---I discarded random subsets of the data until it did not contain
bullshit any more.

\begin{center}
\begin{tabular}{rl}
Rob: &  etsy, nuuog \\
Chris: &  nurm \\
David: &  wafflucinations \\
Lea: &  hnfff \\
Reed: &  pansepticon \\
Jessica: &  gruntle \\
\end{tabular}
\end{center}

From this we can conclude that 16\% of people wish {\it nurm} was a word, and so on.
These words did not come with definitions, except for {\it gruntle}, which Jessica
gives as ``the opposite of distgruntle''. This is actually already a word, but
it was the inspiration for Section~\ref{sec:backformation}. {\it etsy} is the name
of a popular on-line crafts community so I don't know why Rob would suggest that.
The meaning of {\it wafflucinations} is clear from morphological analysis.

% TODO: length normalization. just take the average (or max?)
% probability for a word of length n, and normalize by that (which means what?)?

@@

\section{Conclusion}

\subsection{Recommendations}

@@

In this paper I investigated several different ways of answering the
question: What words ought to exist? Each method produces different
words, and some don't work that well, but nonetheless we have several
rich sources of words, each time scientifically justified.

Nobody really likes writing conclusions, so I will just now summarize
the paper by the giving the most likely sentences that I have not
already written.

% XXXX
